R for Clinical Study Reports and Submission

### 2021-09-16 tlf-ae-spec.Rmd
fix typo in ae tables

For illustration purpose, we only provide count in the simplified table. 
The percentage of participants 
for each AE criteria can be calculated as in Chapter \@ref(aesummary).

In this way, let's focus on the analysis script for advanced feature for table layout.

- group content: AE can be summarized in multiple nested layer. 
(e.g. by system organ class (SOC, `AESOC`) and specific AE term (`AEDECOD`))
- pagenization: there are many AE terms that can not be covered in one page. 
Column headers and SOC information needs to be repeated in every page.

@@

t1 %>% head(4)
```
In the code below, we count number of subjects in each AE term by SOC, AE term, and treatment group.
Here we also create a new variable `order` and set it as `1` for these counted numbers.

@@

t2 %>% head(4)
```
We prepare reporting data for AE information.

@@

t_ae %>% head(4)
```
We prepare reporting data for analysis population.

@@

t_pop
```
The final report data is saved in `tbl_ae_spec`.
We also add a blank row between population and AE information in the reporting table.

@@

tbl_ae_spec %>% head(4)
```
We start to define the format of the output.

To obtain the nested layout, we use `page_by` argument in `rtf_body` function.
By defining `page_by="AESOC"`, 
`r2rtf` recognize the variable as a group indicator.

After setting `pageby_row = "first_row"`, the first row is displayed as group header. If a group of information is breaked into multiple page, 
the group header row is repeated in each page by default.

We can also customize the text format by providing a matrix that have 
same dimension of the input dataset (i.e. `tbl_ae_spec`). 
In the code below, we illustrate how to display **bold** text for group headers
to highlight the nested structure of the table layout.

@@

More discussion on `page_by`, `group_by` and `subline_by` 
features can be found in the 
[`r2rtf` package website](https://merck.github.io/r2rtf/articles/example-sublineby-pageby-groupby.html.

@@

In conclusion, the procedure to generate a specific AE summary table as shown in the above example is listed as follows:

- Step 1: Read data into R, i.e., `adae` and `adsl`.
- Step 2: Count the number of subjects by SOC and treatment group (rows with bold text) and store it in a data frame `t1`.
- Step 3: Count the number of subjects in each AE term by SOC, AE term, and treatment group (rows without bold text) and store it in a data frame `t2`.
- Step 4: Rowly bind `t1` and `t2` into `t_ae`.
- Step 5: Count the number of subjects in each arm as `t_pop`.
- Step 6: Rowly bind `t_pop` with `t_ae` as `tbl_ae_spec`.
- Step 7: Format the output by `r2rtf`.

>>>>

For illustration purposes, we only provide count in the simplified
table. The percentage of participants for each AE criteria can be
calculated as in Chapter \@ref(aesummary).

In this way, let's focus on the analysis script for two advanced
features for table layout.

-   group content: AE can be summarized in multiple nested layers. (e.g.
    by system organ class (SOC, `AESOC`) and specific AE term
    (`AEDECOD`))
-   pagenization: there are many AE terms that can not be covered in one
    page. Column headers and SOC information need to be repeated on
    every page.

@@

t1 %>% head(4)
```

In the code below, we count the number of subjects in each AE term by
SOC, AE term, and treatment group. Here we also create a new variable
`order` and set it as `1` for these counted numbers.

@@

t2 %>% head(4)
```

We prepare reporting data for AE information.

@@

t_ae %>% head(4)
```

We prepare reporting data for analysis population.

@@

t_pop
```

The final report data is saved in `tbl_ae_spec`. We also add a blank row
between population and AE information in the reporting table.

@@

tbl_ae_spec %>% head(4)
```

We start to define the format of the output.

To obtain the nested layout, we use the `page_by` argument in the
`rtf_body` function. By defining `page_by="AESOC"`, `r2rtf` recognize
the variable as a group indicator.

After setting `pageby_row = "first_row"`, the first row is displayed as
group header. If a group of information is broken into multiple pages,
the group header row is repeated on each page by default.

We can also customize the text format by providing a matrix that has the
same dimension as the input dataset (i.e. `tbl_ae_spec`). In the code
below, we illustrate how to display **bold** text for group headers to
highlight the nested structure of the table layout.

@@

More discussion on `page_by`, `group_by`, and `subline_by` features can
be found in the [`r2rtf` package
website](https://merck.github.io/r2rtf/articles/example-sublineby-pageby-groupby.html).

@@

The procedure to generate a baseline characteristics table can be
summarized as follow:

-   Step 1: Read data into R, i.e., `adae` and `adsl`.
-   Step 2: Count the number of subjects by SOC and treatment group
    (rows with bold text) and save into `t1`.
-   Step 3: Count the number of subjects in each AE term by SOC, AE
    term, and treatment group (rows without bold text) and save into
    `t2`.
-   Step 4: Bind `t1` and `t2` by row into `t_ae`.
-   Step 5: Count the number of subjects in each arm as `t_pop`.
-   Step 6: Bind `t_pop` and `t_ae` by row into `tbl_ae_spec`.
-   Step 7: Format the output by `r2rtf`.

### 2021-09-16 tlf-ae-summary.Rmd
fix typo in ae tables

- With one or more adverse events
- With drug-related adverse events
- With serious adverse events
- With serious drug-related adverse events
- Who died"

@@

tidy_ae %>% head(4)
```
We summarize the number and percentage of patients meet each AE criteria.

@@

ana %>% head(4)
```
We prepare reporting ready dataset for each AE criteria.

@@

We prepare reporting ready dataset for analysis population.

@@

t_pop
```
The final report data is saved in `tbl_ae_summary`

@@

In conclusion, the procedure to generate a AE summary table as shown in the above example is listed as follows:

>>>>

-   With one or more adverse events
-   With drug-related adverse events
-   With serious adverse events
-   With serious drug-related adverse events
-   Who "died"

@@

tidy_ae %>% head(4)
```

We summarize the number and percentage of patients who meet each AE criteria.

@@

ana %>% head(4)
```

We prepare reporting-ready dataset for each AE criteria.

@@

We prepare reporting-ready dataset for the analysis population.

@@

t_pop
```

The final report data is saved in `tbl_ae_summary`.

@@

The procedure to generate a AE summary table can be summarized as follow：
